%----------------------------------------------------------------------------------------
%    BIBLIOGRAPHY
%----------------------------------------------------------------------------------------

\label{Bibliography}
\lhead{\emph{Bibliography}}

\begin{thebibliography}{55}

\bibitem{topol2019}
Tourkiah Alessa, Mark S. Hawley, Emma S. Hock, and Luc de Witte. 2019. \emph{Smartphone Apps to Support Self-Management of Hypertension: Review and Content Analysis}. JMIR Mhealth Uhealth 7, 5 (28 May 2019), e13645.

\bibitem{ayers2023}
John W. Ayers, Adam Poliak, Mark Dredze, Eric C. Leas, Zechariah Zhu, Jessica B. Kelley, Dennis J. Faix, Aaron M. Goodman, Christopher A. Longhurst, Michael Hogarth, and Davey M. Smith. 2023. \emph{Comparing Physician and Artificial Intelligence Chatbot Responses to Patient Questions Posted to a Public Social Media Forum}. JAMA Internal Medicine (2023).

\bibitem{bigini2023}
Gioele Bigini, Mirko Zichichi, Emanuele Lattanzi, Stefano Ferretti, and Gabriele D’Angelo. 2023. \emph{On the Decentralization of Health Systems for Data Availability: A DLT-based Architecture}. In 2023 IEEE 20th Consumer Communications \& Networking Conference (CCNC), 372--377.

\bibitem{cascella2023}
Marco Cascella, Jonathan Montomoli, Valentina Bellini, and Elena Bignami. 2023. \emph{Evaluating the Feasibility of ChatGPT in Healthcare: An Analysis of Multiple Clinical and Research Scenarios}. Journal of Medical Systems 47, 1 (2023), 33.

\bibitem{clark1991}
Noreen M. Clark, Marshall H. Becker, Nancy K. Janz, Kate Lorig, William Rakowski, and Lynda Anderson. 1991. \emph{Self-Management of Chronic Disease by Older Adults: A Review and Questions for Research}. Journal of Aging and Health 3, 1 (1991), 3--27.

\bibitem{gbd2020}
GBD 2019 Risk Factors Collaborators. 2020. \emph{Global Burden of 87 Risk Factors in 204 Countries and Territories, 1990--2019: A Systematic Analysis for the Global Burden of Disease Study 2019}. The Lancet 396, 10258 (2020), 1223--1249.

\bibitem{dineen2019}
Sarah Dineen-Griffin, Victoria Garcia-Cardenas, Kylie Williams, and Shalom I. Benrimoj. 2019. \emph{Helping Patients Help Themselves: A Systematic Review of Self-Management Support Strategies in Primary Health Care Practice}. PLOS ONE 14, 8 (2019), 1--29.

\bibitem{echeazarra2021}
Leyre Echeazarra, Juanan Pereira, and Ramón Saracho. 2021. \emph{TensioBot: A Chatbot Assistant for Self-Managed in-House Blood Pressure Checking}. Journal of Medical Systems 45, 4 (2021), 54. \url{https://doi.org/10.1007/s10916-021-01730-x}.

\bibitem{editorial2023}
Editorial. 2023. \emph{Will ChatGPT Transform Healthcare?} Nature Medicine 29, 3 (2023), 505--506.

\bibitem{parati2021}
Gianfranco Parati et al. 2021. \emph{Home Blood Pressure Monitoring: Methodology, Clinical Relevance and Practical Application: A 2021 Position Paper by the Working Group on Blood Pressure Monitoring and Cardiovascular Variability of the European Society of Hypertension}. Journal of Hypertension 39, 9 (2021).

\bibitem{vaishya2023}
Raju Vaishya et al. 2023. \emph{ChatGPT: Is This Version Good for Healthcare and Research?} Diabetes \& Metabolic Syndrome: Clinical Research \& Reviews 17, 4 (2023), 102744.

\bibitem{eu2016}
European Parliament and Council of the European Union. 2016. \emph{Regulation (EU) 2016/679: General Data Protection Regulation (GDPR)}. Official Journal of the European Union. \url{https://eur-lex.europa.eu/legal-content/EN/TXT/?uri=CELEX:32016R0679}.

\bibitem{heidel2020}
Alexandra Heidel and Christian Hagist. 2020. \emph{Potential Benefits and Risks Resulting from the Introduction of Health Apps and Wearables Into the German Statutory Health Care System: Scoping Review}. JMIR Mhealth Uhealth 8, 9 (23 Sep 2020), e16444. \url{https://doi.org/10.2196/16444}.

\bibitem{hernandez2021}
N. Hernandez, L. Castro, J. Medina-Quero, J. Favela, L. Michan, and W. Ben. Mortenson. 2021. \emph{Scoping Review of Healthcare Literature on Mobile, Wearable, and Textile Sensing Technology for Continuous Monitoring}. Journal of Healthcare Informatics Research 5, 3 (2021), 270--299. \url{https://doi.org/10.1007/s41666-020-00087-z}.

\bibitem{koopman2020}
Richelle J. Koopman, Shannon M. Canfield, Jeffery L. Belden, Pete Wegier, Victoria A. Shaffer, K. D. Valentine, Akshay Jain, Linsey M. Steege, Sonal J. Patil, Mihail Popescu, and Michael L. LeFevre. 2020. \emph{Home Blood Pressure Data Visualization for the Management of Hypertension: Designing for Patient and Physician Information Needs}. BMC Medical Informatics and Decision Making 20, 1 (2020), 195. \url{https://doi.org/10.1186/s12911-020-01194-y}.

\bibitem{laranjo2018}
Liliana Laranjo, Adam G. Dunn, Huong Ly Tong, Ahmet Baki Kocaballi, Jessica Chen, Rabia Bashir, Didi Surian, Blanca Gallego, Farah Magrabi, Annie Y. S. Lau, and Enrico Coiera. 2018. \emph{Conversational Agents in Healthcare: A Systematic Review}. Journal of the American Medical Informatics Association 25, 9 (2018), 1248--1258.

\bibitem{mohammed2014}
Niloofar Mohammadzadeh and Reza Safdari. 2014. \emph{Patient Monitoring in Mobile Health: Opportunities and Challenges}. Medical Archives 68, 1 (2014), 57--60. \url{https://doi.org/10.5455/medarh.2014.68.57-60}.

\bibitem{montagna2023}
Sara Montagna, Stefano Mariani, and Martino Francesco Pengo. 2023. \emph{A Chatbot-based Recommendation Framework for Hypertensive Patients}. In 2023 IEEE 36th International Symposium on Computer-Based Medical Systems (CBMS), 730--733.

\bibitem{hipaa1996}
Act of Congress enacted by the 104th United States Congress. 1996. \emph{Health Insurance Portability and Accountability Act of 1996 (HIPAA)}. Public law 104-191.

\bibitem{akker2015}
Harm op den Akker, Miriam Cabrita, Rieks op den Akker, Valerie M. Jones, and Hermie J. Hermens. 2015. \emph{Tailored Motivational Message Generation: A Model and Practical Framework for Real-Time Physical Activity Coaching}. Journal of Biomedical Informatics 55 (2015), 104--115. \url{https://doi.org/10.1016/j.jbi.2015.03.005}.

\bibitem{pan2020}
Xudong Pan, Mi Zhang, Shouling Ji, and Min Yang. 2020. \emph{Privacy Risks of General-Purpose Language Models}. In 2020 IEEE Symposium on Security and Privacy (SP), 1314--1331.

\bibitem{qiu2023}
Jianing Qiu, Lin Li, Jiankai Sun, Jiachuan Peng, Peilun Shi, Ruiyang Zhang, Yinzhao Dong, Kyle Lam, Frank P. W. Lo, Bo Xiao, Wu Yuan, Dong Xu, and Benny Lo. 2023. \emph{Large AI Models in Health Informatics: Applications, Challenges, and the Future}. arXiv:2303.11568, 2023.

\bibitem{tabak2014}
Monique Tabak, Harm op den Akker, and Hermie Hermens. 2014. \emph{Motivational Cues as Real-Time Feedback for Changing Daily Activity Behavior of Patients with COPD}. Patient Education and Counseling 94, 3 (2014), 372--378.

\bibitem{zichichi2020a}
Mirko Zichichi, Stefano Ferretti, Gabriele D’Angelo, and Víctor Rodríguez-Doncel. 2020. \emph{Personal Data Access Control Through Distributed Authorization}. In 2020 IEEE 19th International Symposium on Network Computing and Applications (NCA), 1--4.

\bibitem{zichichi2020b}
Mirko Zichichi, Stefano Ferretti, and Gabriele D’Angelo. 2020. \emph{On the Efficiency of Decentralized File Storage for Personal Information Management Systems}. In 2020 IEEE Symposium on Computers and Communications (ISCC), 1--6.

\bibitem{zichichi2022}
Mirko Zichichi, Stefano Ferretti, Gabriele D’Angelo, and Víctor Rodríguez-Doncel. 2022. \emph{Data Governance through a Multi-DLT Architecture in View of the GDPR}. Cluster Computing 25(6), 4515--4542.

\bibitem{gpt4tech}
OpenAI et al. 2023. \emph{GPT-4 Technical Report}. arXiv:2303.08774, 2023.

\bibitem{llama2023}
H. Touvron, T. Lavril, G. Izacard, X. Martinet, M.-A. Lachaux, T. Lacroix, B. Rozière, N. Goyal, E. Hambro, F. Azhar, A. Rodriguez, A. Joulin, E. Grave, and G. Lample. 2023. \emph{Llama: Open and Efficient Foundation Language Models}. arXiv:2302.13971, 2023.

\bibitem{llama2}
H. Touvron, L. Martin, K. Stone, P. Albert, A. Almahairi, Y. Babaei, N. Bashlykov, S. Batra, P. Bhargava, S. Bhosale, et al. 2023. \emph{Llama 2: Open Foundation and Fine-Tuned Chat Models}. arXiv:2307.09288, 2023.

\bibitem{mistral7b}
A. Q. Jiang, A. Sablayrolles, A. Mensch, C. Bamford, D. S. Chaplot, D. de las Casas, F. Bressand, G. Lengyel, G. Lample, L. Saulnier, L. R. Lavaud, M.-A. Lachaux, P. Stock, T. L. Scao, T. Lavril, T. Wang, T. Lacroix, and W. E. Sayed. 2023. \emph{Mistral 7b}. arXiv:2310.06825, 2023.

\bibitem{montagna2023b}
Sara Montagna, Stefano Ferretti, Lorenz Cuno Klopfenstein, Antonio Florio, and Martino Pengo. 2023. \emph{Data Decentralisation of LLM-Based Chatbot Systems in Chronic Disease Self-Management}. In Proceedings of the ACM International Conference on Information Technology for Social Good (GoodIT '23), 1--8. doi:10.1145/3582515.3609536.

\bibitem{wen2023b}
Bo Wen, Raquel Norel, Julia Liu, Thaddeus Stappenbeck, Farhana Zulkernine, and Huamin Chen. 2023. \emph{Leveraging Large Language Models for Patient Engagement: The Power of Conversational AI in Digital Health}. In Proceedings of the IEEE International Conference on Digital Health (ICDH), 1--8. doi:10.1109/ICDH.2023.00001.

\bibitem{dhurandhar2024}
A. Dhurandhar, R. Nair, M. Singh, E. Daly, and K. N. Ramamurthy. 2024. \emph{Ranking Large Language Models without Ground Truth}. arXiv:2402.14860, 2024.

\bibitem{ueda2024}
D. Ueda, T. Kakinuma, S. Fujita, K. Kamagata, Y. Fushimi, R. Ito, Y. Matsui, T. Nozaki, T. Nakaura, N. Fujima, et al. 2024. \emph{Fairness of Artificial Intelligence in Healthcare: Review and Recommendations}. Japanese Journal of Radiology, 42(1), 3--15.

\bibitem{kiseleva2022}
A. Kiseleva and D. Kotzinos. 2022. \emph{Transparency of AI in Healthcare as a Multilayered System of Accountabilities: Between Legal Requirements and Technical Limitations}. Frontiers in Artificial Intelligence, 5. Retrieved from \url{https://www.frontiersin.org/articles/10.3389/frai.2022.879603/full}.

\bibitem{harding2023}
T. Harding, E. Cunningham, and A. Jebran. 2023. \emph{Navigating the Regulatory Landscape: An Analysis of Legal and Ethical Oversight for Large Language Models (LLMs)}. Retrieved from \url{https://www.himss.org/resources/section-5-navigating-regulatory-landscape-analysis-legal-and-ethical-oversight-large}.

\bibitem{park2024}
Y.-J. Park, A. Pillai, J. Deng, E. Guo, M. Gupta, M. Paget, and C. Naugler. 2024. \emph{Assessing the Research Landscape and Clinical Utility of Large Language Models: A Scoping Review}. BMC Medical Informatics and Decision Making, 24(1), 72.

\bibitem{ullah2024}
E. Ullah, A. Parwani, M. M. Baig, and R. Singh. 2024. \emph{Challenges and Barriers of Using Large Language Models (LLM) such as ChatGPT for Diagnostic Medicine with a Focus on Digital Pathology--A Recent Scoping Review}. Diagnostic Pathology, 19(43). Retrieved from \url{https://diagnosticpathology.biomedcentral.com/articles/10.1186/s13000-024-01464-7}.

\bibitem{mumtaz2023}
H. Mumtaz, M. H. Riaz, H. Wajid, M. Saqib, M. H. Zeeshan, S. E. Khan, Y. R. Chauhan, H. Sohail, and L. I. Vohra. 2023. \emph{Current Challenges and Potential Solutions to the Use of Digital Health Technologies in Evidence Generation: A Narrative Review}. Frontiers in Digital Health, 5, 1203945.

\end{thebibliography}

\end{document}
