%----------------------------------------------------------------------------------------
%    BIBLIOGRAPHY
%----------------------------------------------------------------------------------------

\label{Bibliography}
\lhead{\emph{Bibliography}}

\begin{thebibliography}{55}

\bibitem{topol2019}
Tourkiah Alessa, Mark S. Hawley, Emma S. Hock, and Luc de Witte. 2019. \emph{Smartphone Apps to Support Self-Management of Hypertension: Review and Content Analysis}. JMIR Mhealth Uhealth 7, 5 (28 May 2019), e13645.

\bibitem{ayers2023}
John W. Ayers, Adam Poliak, Mark Dredze, Eric C. Leas, Zechariah Zhu, Jessica B. Kelley, Dennis J. Faix, Aaron M. Goodman, Christopher A. Longhurst, Michael Hogarth, and Davey M. Smith. 2023. \emph{Comparing Physician and Artificial Intelligence Chatbot Responses to Patient Questions Posted to a Public Social Media Forum}. JAMA Internal Medicine (2023).

\bibitem{bigini2023}
Gioele Bigini, Mirko Zichichi, Emanuele Lattanzi, Stefano Ferretti, and Gabriele D’Angelo. 2023. \emph{On the Decentralization of Health Systems for Data Availability: A DLT-based Architecture}. In 2023 IEEE 20th Consumer Communications \& Networking Conference (CCNC), 372--377.

\bibitem{cascella2023}
Marco Cascella, Jonathan Montomoli, Valentina Bellini, and Elena Bignami. 2023. \emph{Evaluating the Feasibility of ChatGPT in Healthcare: An Analysis of Multiple Clinical and Research Scenarios}. Journal of Medical Systems 47, 1 (2023), 33.

\bibitem{clark1991}
Noreen M. Clark, Marshall H. Becker, Nancy K. Janz, Kate Lorig, William Rakowski, and Lynda Anderson. 1991. \emph{Self-Management of Chronic Disease by Older Adults: A Review and Questions for Research}. Journal of Aging and Health 3, 1 (1991), 3--27.

\bibitem{dineen2019}
Sarah Dineen-Griffin, Victoria Garcia-Cardenas, Kylie Williams, and Shalom I. Benrimoj. 2019. \emph{Helping Patients Help Themselves: A Systematic Review of Self-Management Support Strategies in Primary Health Care Practice}. PLOS ONE 14, 8 (2019), 1--29.

\bibitem{editorial2023}
Editorial. 2023. \emph{Will ChatGPT Transform Healthcare?} Nature Medicine 29, 3 (2023), 505--506.

\bibitem{vaishya2023}
Raju Vaishya et al. 2023. \emph{ChatGPT: Is This Version Good for Healthcare and Research?} Diabetes \& Metabolic Syndrome: Clinical Research \& Reviews 17, 4 (2023), 102744.

\bibitem{eu2016}
European Parliament and Council of the European Union. 2016. \emph{Regulation (EU) 2016/679: General Data Protection Regulation (GDPR)}. Official Journal of the European Union. \url{https://eur-lex.europa.eu/legal-content/EN/TXT/?uri=CELEX:32016R0679}.

\bibitem{hernandez2021}
N. Hernandez, L. Castro, J. Medina-Quero, J. Favela, L. Michan, and W. Ben. Mortenson. 2021. \emph{Scoping Review of Healthcare Literature on Mobile, Wearable, and Textile Sensing Technology for Continuous Monitoring}. Journal of Healthcare Informatics Research 5, 3 (2021), 270--299. \url{https://doi.org/10.1007/s41666-020-00087-z}.

\bibitem{koopman2020}
Richelle J. Koopman, Shannon M. Canfield, Jeffery L. Belden, Pete Wegier, Victoria A. Shaffer, K. D. Valentine, Akshay Jain, Linsey M. Steege, Sonal J. Patil, Mihail Popescu, and Michael L. LeFevre. 2020. \emph{Home Blood Pressure Data Visualization for the Management of Hypertension: Designing for Patient and Physician Information Needs}. BMC Medical Informatics and Decision Making 20, 1 (2020), 195. \url{https://doi.org/10.1186/s12911-020-01194-y}.

\bibitem{montagna2023}
Sara Montagna, Stefano Mariani, and Martino Francesco Pengo. 2023. \emph{A Chatbot-based Recommendation Framework for Hypertensive Patients}. In 2023 IEEE 36th International Symposium on Computer-Based Medical Systems (CBMS), 730--733.

\bibitem{hipaa1996}
Act of Congress enacted by the 104th United States Congress. 1996. \emph{Health Insurance Portability and Accountability Act of 1996 (HIPAA)}. Public law 104-191.

\bibitem{pan2020}
Xudong Pan, Mi Zhang, Shouling Ji, and Min Yang. 2020. \emph{Privacy Risks of General-Purpose Language Models}. In 2020 IEEE Symposium on Security and Privacy (SP), 1314--1331.

\bibitem{qiu2023}
Jianing Qiu, Lin Li, Jiankai Sun, Jiachuan Peng, Peilun Shi, Ruiyang Zhang, Yinzhao Dong, Kyle Lam, Frank P. W. Lo, Bo Xiao, Wu Yuan, Dong Xu, and Benny Lo. 2023. \emph{Large AI Models in Health Informatics: Applications, Challenges, and the Future}. arXiv:2303.11568, 2023.

\bibitem{tabak2014}
Monique Tabak, Harm op den Akker, and Hermie Hermens. 2014. \emph{Motivational Cues as Real-Time Feedback for Changing Daily Activity Behavior of Patients with COPD}. Patient Education and Counseling 94, 3 (2014), 372--378.

\bibitem{zichichi2020a}
Mirko Zichichi, Stefano Ferretti, Gabriele D’Angelo, and Víctor Rodríguez-Doncel. 2020. \emph{Personal Data Access Control Through Distributed Authorization}. In 2020 IEEE 19th International Symposium on Network Computing and Applications (NCA), 1--4.

\bibitem{zichichi2020b}
Mirko Zichichi, Stefano Ferretti, and Gabriele D’Angelo. 2020. \emph{On the Efficiency of Decentralized File Storage for Personal Information Management Systems}. In 2020 IEEE Symposium on Computers and Communications (ISCC), 1--6.

\bibitem{zichichi2022}
Mirko Zichichi, Stefano Ferretti, Gabriele D’Angelo, and Víctor Rodríguez-Doncel. 2022. \emph{Data Governance through a Multi-DLT Architecture in View of the GDPR}. Cluster Computing 25(6), 4515--4542.

\bibitem{gpt4tech}
OpenAI et al. 2023. \emph{GPT-4 Technical Report}. arXiv:2303.08774, 2023.

\bibitem{llama2023}
H. Touvron, T. Lavril, G. Izacard, X. Martinet, M.-A. Lachaux, T. Lacroix, B. Rozière, N. Goyal, E. Hambro, F. Azhar, A. Rodriguez, A. Joulin, E. Grave, and G. Lample. 2023. \emph{Llama: Open and Efficient Foundation Language Models}. arXiv:2302.13971, 2023.

\bibitem{mistral7b}
A. Q. Jiang, A. Sablayrolles, A. Mensch, C. Bamford, D. S. Chaplot, D. de las Casas, F. Bressand, G. Lengyel, G. Lample, L. Saulnier, L. R. Lavaud, M.-A. Lachaux, P. Stock, T. L. Scao, T. Lavril, T. Wang, T. Lacroix, and W. E. Sayed. 2023. \emph{Mistral 7b}. arXiv:2310.06825, 2023.

\bibitem{wen2023b}
Bo Wen, Raquel Norel, Julia Liu, Thaddeus Stappenbeck, Farhana Zulkernine, and Huamin Chen. 2023. \emph{Leveraging Large Language Models for Patient Engagement: The Power of Conversational AI in Digital Health}. In Proceedings of the IEEE International Conference on Digital Health (ICDH), 1--8. doi:10.1109/ICDH.2023.00001.

\bibitem{dhurandhar2024}
A. Dhurandhar, R. Nair, M. Singh, E. Daly, and K. N. Ramamurthy. 2024. \emph{Ranking Large Language Models without Ground Truth}. arXiv:2402.14860, 2024.

\bibitem{kiseleva2022}
A. Kiseleva and D. Kotzinos. 2022. \emph{Transparency of AI in Healthcare as a Multilayered System of Accountabilities: Between Legal Requirements and Technical Limitations}. Frontiers in Artificial Intelligence, 5. Retrieved from \url{https://www.frontiersin.org/articles/10.3389/frai.2022.879603/full}.


\bibitem{park2024}
Y.-J. Park, A. Pillai, J. Deng, E. Guo, M. Gupta, M. Paget, and C. Naugler. 2024. \emph{Assessing the Research Landscape and Clinical Utility of Large Language Models: A Scoping Review}. BMC Medical Informatics and Decision Making, 24(1), 72.

\bibitem{sqlite3}
Python Software Foundation. (2024). \emph{sqlite3 — DB-API 2.0 interface for SQLite databases}. \url{https://docs.python.org/3/library/sqlite3.html}

\bibitem{httpcore}
Encode. (2024). \emph{httpcore Documentation}. \url{https://httpcore.readthedocs.io/en/latest/}

\bibitem{selenium}
Selenium. (2024). \emph{Selenium Documentation}. \url{https://www.selenium.dev/documentation/}

\bibitem{beautifulsoup}
Richardson, L. (2024). \emph{BeautifulSoup Documentation}. \url{https://www.crummy.com/software/BeautifulSoup/bs4/doc/}

\bibitem{time}
Python Software Foundation. (2024). \emph{time — Time access and conversions}. \url{https://docs.python.org/3/library/time.html}

\bibitem{json}
Python Software Foundation. (2024). \emph{json — JSON encoder and decoder}. \url{https://docs.python.org/3/library/json.html}

\bibitem{chainlit}
Chainlit Documentation. (2024). \emph{Chainlit - Overview}. \url{https://docs.chainlit.io/get-started/overview}

\bibitem{pinecone}
Pinecone Documentation. (2024). \emph{Pinecone Vector Database Documentation}. \url{https://docs.pinecone.io/reference/python-sdk}

\bibitem{langchain}
LangChain. (2024). \emph{LangChain Documentation}. \url{https://python.langchain.com/docs/introduction/}

\bibitem{dotenv}
Python Dotenv. (2024). \emph{python-dotenv Documentation}. \url{https://pypi.org/project/python-dotenv/}

\bibitem{sentence_transformers}
Reimers, N., \& Gurevych, I. (2024). \emph{SentenceTransformers Documentation}. \url{https://www.sbert.net/}

\bibitem{twilio}
Twilio. (2024). \emph{Twilio SMS Python Quickstart}. \url{https://www.twilio.com/docs/messaging/quickstart/python}

\bibitem{os}
Python Software Foundation. (2024). \emph{os — Operating system interfaces}. \url{https://docs.python.org/3/library/os.html}

\bibitem{logging}
Python Software Foundation. (2024). \emph{logging — Logging facility}. \url{https://docs.python.org/3/library/logging.html}

\bibitem{warnings}
Python Software Foundation. (2024). \emph{warnings — Warning control}. \url{https://docs.python.org/3/library/warnings.html}

\bibitem{asyncio}
Python Software Foundation. (2024). \emph{asyncio — Asynchronous I/O}. \url{https://docs.python.org/3/library/asyncio.html}

\bibitem{flask}
Pallets. (2024). \emph{Flask Documentation}. \url{https://flask.palletsprojects.com/en/latest/}

\bibitem{subprocess}
Python Software Foundation. (2024). \emph{subprocess — Subprocess management}. \url{https://docs.python.org/3/library/subprocess.html}

\bibitem{threading}
Python Software Foundation. (2024). \emph{threading — Thread-based parallelism}. \url{https://docs.python.org/3/library/threading.html}

\bibitem{requests}
Reitz, K. (2024). \emph{Requests: HTTP for Humans}. \url{https://docs.python-requests.org/en/latest/}

\bibitem{numpy}
NumPy Developers. (2024). \emph{NumPy Documentation}. \url{https://numpy.org/doc/}

\end{thebibliography}

\end{document}
